Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-09-17 second_submit/arcticgrids_trackchanges.tex
second submit complete

\begin{keypoints}
\item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two Arctic refined mesh configurations.
\item Quasi-uniform unstructured grids degrade the Greenland Ice Sheet mass balance compared to latitude-longitude grids, at the conventional $1^{\circ}$ resolution.
\item The refined Arctic meshes substantially improve the surface mass balance over conventional grid resolutions.
\end{keypoints}

@@

Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS). The finite-volume dynamical core uses structured, latitude-longitude grids, whereas the spectral-element dynamical core is built on unstructured meshes, permitting grid flexibility such as quasi-uniform grid spacing globally. The $1^{\circ}-2^{\circ}$ latitude-longitude and quasi-uniform unstructured grids systematically overestimate both accumulation and ablation over the GrIS. Of these $1^{\circ}-2^{\circ}$ grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids because they have more degrees of freedom to represent the GrIS. Two Arctic-refined meshes, with $1/4^{\circ}$ and $1/8^{\circ}$ refinement over Greenland, were developed for the spectral-element dynamical core and are documented here as newly supported configurations in CESM2.2. The Arctic meshes substantially improve the simulated clouds and precipitation rates in the Arctic. Over Greenland, these meshes skillfully represent accumulation and ablation processes, leading to a more realistic GrIS surface mass balance. As CESM is in the process of transitioning away from conventional latitude-longitude grids, these new Arctic-refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids, albeit at increased computational cost.

@@

The mass balance of Earth's big ice sheets is crucially important for understanding controls on global sea-level rise. However, the scale of the processes needed to represent ice sheet mass balance is challenging to resolve in conventional Earth System Models. This study evaluates the ability of different grids and atmospheric solvers (i.e, the dynamical core) in CESM2.2 to resolve the surface mass balance (SMB) of the Greenland Ice Sheet. We show that the ongoing transition away from latitude-longitude grids, towards quasi-uniform unstructured grids in CESM2.2 leads to a degradation of the simulated SMB. Two variable-resolution grids with enhanced resolution over Greenland are developed and incorporated into the release of CESM2.2, which substantially improves the SMB over the latitude longitude grids.
%[ enter your Plain Language Summary here or delete this section]

@@

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change. GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a, hereafter referred to as \textit{lat-lon} grids). Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain. With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost (hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic due to the polar singularity). In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids (i.e., grids with enhanced resolution over a particular region), to understand their impacts on the simulated Arctic climate. We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.

@@

Dycores built on lat-lon grids have some advantages over dycores built on unstructured grids. Lat-lon coordinate lines are orthogonal, and aligned with zonally symmetric circulations that characterize many large-scale features of Earth's atmosphere. \citeA{LJTN2010JAMES} has experimented with rotating lat-lon models such that their coordinate lines no longer align with an idealized, zonally balanced circulation. For the finite-volume lat-lon dycore considered in this paper (hereafter \textit{FV}), numerical errors were shown to be largest when the polar singularity is rotated into the baroclinic zone ($45^{\circ}$N latitude), generating spurious wave growth much earlier in the simulation than for other rotation angles. This illustrates the advantages of coordinate surfaces aligned with latitude bands, albeit an extreme example where the polar singularity and the polar filter are also contributing to the spurious wave growth. The unstructured grids all generate spurious baroclinic waves earlier in the simulations than the (unrotated) lat-lon models, although the unstructured model considered in this paper, the spectral-element dycore (hereafter \textit{SE}), holds a balanced zonal flow without spurious wave growth appreciably longer than the rotated FV experiments \cite{LJTN2010JAMES}. And unlike FV, the SE dycore has the same error characteristics regardless of how the grid is rotated.

The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which has a large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}. Unstructured grids support quasi-uniform grid spacing globally, and there is no pole problem (e.g., Figure~\ref{fig:uni-grids}c). This is in part why unstructured grids are becoming more common; their improved performance on massively parallel systems and lack of constraints on grid structure \cite{TTI1997JCP,PL2007JCP,WETAL2013GMD}. This increased grid flexibility allows for the adoption of variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement. In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).  However, local grid refinement comes at the cost of a smaller CFL-limited time step in the refined region; the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.

We emphasize that the pole problem is a distinctive feature of the dycore in atmospheric models. Polar filters do not directly interfere with the physical parameterizations, nor do they have any bearing on the surface models; e.g., the land model can take full advantage of the greater number of grid cells in polar regions on lat-lon grids. This is particularly relevant for the surface mass balance (\textit{SMB}; the integrated sum of precipitation and runoff) of the Greenland Ice Sheet, which relies on hydrological processes represented in the land model.

The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales \cite{Fyke_EA_2018_Rev_Geophys} that are difficult to represent in GCMs \cite{P2010CC}. GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation. The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias in the interior \cite{P2000GPC,VETAL2018TC}. GrIS ablation areas (marginal regions where the annual SMB is negative) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation. These narrow ablation zones are not fully resolved in low-resolution GCMs, and may further degrade the simulated SMB. More recently, GCMs such as CESM and the NASA Goddard Institute for Space Studies GCM \cite{AETAL2019JGR}, have implemented an elevation class downscaling scheme for computing the SMB. The downscaling helps to resolve these narrow ablation zones in GCMs \cite{SETAL2019TC}, but large SMB biases remain. For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.

@@

In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} used the VR capabilities of the SE dycore in CESM, generating grids where Greenland is represented with up to $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution. The simulated SMB compared favorably to the SMB from RCMs and observations. The VR approach is therefore emerging as a powerful tool for simulating and understanding the GrIS and its response to different forcing scenarios.

@@

CESM2.2 is a CMIP6-class \cite<Coupled Model Intercomparison Project Phase 6;>{CMIP6} Earth System Model maintained by the National Center for Atmosphere Research. CESM2.2 contains sub-component models for the atmosphere, land, ocean, sea-ice, and land-ice, and can be configured to run with varying degrees of complexity. All simulations described in this study use an identical transient 1979-1998 Atmospheric Model Intercomparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}. In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2. The land-ice component is not dynamically active in $\tt FHIST$. However, the surface mass balance is computed by the land model before being passed to the land-ice component; $\tt FHIST$ includes the necessary functionality to simulate the SMB of the Greenland Ice Sheet.

@@

The FV dycore integrates the hydrostatic equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}. The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}. Hyperviscous damping is applied to the divergent modes, and is increased in the top few layers (referred to as a \textit{sponge layer}) to prevent undesirable interactions with the model top, such as wave reflection \cite{L2011IJHPC}. A polar filter damps computational instability due to the convergence of meridians, permitting a longer time step. It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the meridional direction \cite{ST1995GEOS}. The form of the filter is designed to slow down the propagation of large zonal wave-numbers to satisfy the CFL condition of the shortest resolved wave at some reference latitude.

@@

The SE dycore integrates the hydrostatic equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}. The computational domain is a cubed-sphere grid (obtained by projecting each face of a cube onto a sphere) tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}). Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre quadrature points. This results in 16 nodal points per element, with 12 of the points lying on the (shared) element boundary. Communication between elements uses the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries to reconcile overlapping nodal values and produce a continuous global basis set.

@@

SE is a next-generation dycore, and is less mature than the FV dycore due to its shorter history. In CESM2.2, the SE numerics have been enhanced relative to CESM2.0 to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}. Future versions of CESM will likely include further optimizations and enhancements to the (more junior) SE dycore.

The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over quasi-uniform resolution grids. First, as the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. Second, the topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{LetAl2015GMD} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.

When using the SE dycore with quasi-uniform grid spacing, the SE tracer transport scheme is replaced with the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}. Atmospheric tracers have large, nearly discontinuous horizontal gradients that are difficult to represent with spectral methods, which are prone to oscillatory ``Gibbs-ringing" errors \cite{RW1990QJR}. CSLAM has improved tracer property preservation and accelerated multi-tracer transport. It uses a separate grid from the spectral-element dynamics, dividing each element into $3\times3$ control volumes with quasi-equal area. The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the original SE dycore in which the physics are evaluated at Gauss-Lobatto-Legendre points \cite{HL2018MWR}. CSLAM advection is not available in the VR configuration, which instead uses the standard SE tracer transport scheme with the physics evaluated at Gauss-Lobatto-Legendre points.

@@

Three VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic and the United States (Figure~\ref{fig:vr-grids}). This paper serves as the official documentation of these grids. The VR meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region (when using degrees to describe the resolution of unstructured grids, we refer to degrees at the equator). The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution. The $\tt CONUS$ grid contains $1/8^{\circ}$ refinement over the United States, and $1^{\circ}$ everywhere else. The $\tt CONUS$ grid is not discussed any further in this paper; see \citeA{PETAL2020BAMS} for simulations with the $\tt CONUS$ grid.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & cost(25 nodes) & cost(50 nodes) \\ 
   \hline
   $\tt f19$ & FV & 278 & - &1800 & 436.66 & - \\
   $\tt f09$ & FV & 139 & - &1800 & 1534.57 & 2024.24 \\
   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 & 1497.26 & 1683.97 \\
   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 & 1890.48 & 2090.43 \\
   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 & - & - \\
   $\tt Arctic$ & SE & 111 & 28 & 450 & 15947.41 & 16675.45 \\
   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 & 40305.03 & 41036.67 \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{fine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step. FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection. The FV dycore uses lat-lon grids, whereas the SE and SE-CSLAM dycores run on unstructured grids. We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$. The last columns provide the computational costs in core hours per simulated year (CHPSY). The costs are from single-month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}
 \label{tbl:table1}
 \end{table*}

The accuracy of the simulated surface mass balance is expected to be sensitive to grid resolution. Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids, as well as two grids pertaining to the Regional Atmospheric Climate Model \cite<RACMO;>{NETAL2018TC,NETAL2019SCIENCE}, which are used for validation (Table~\ref{tbl:table2}). The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160$ km, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6$ km, similar to the $11$ km grid spacing of the RACMO2.3 grid. The $\tt ne30pg3$ grid has an average $\Delta x=111.2$ km, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60$ km. Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs (Table~\ref{tbl:table1}), the convergence of meridians on the FV grid enhances the resolution over the GrIS. The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8$ km.

@@

\subsection{Computational costs}

The last columns of Table~\ref{tbl:table1} provide cost estimates for the different grids and dycores. The costs, expressed as core hours per simulated year (CHPSY), are taken from single-month runs of $\tt FHIST$, with no i/o, and using 25 nodes (900 tasks) and 50 nodes (1800 tasks) on the Cheyenne supercomputer \cite{cheyenne}. For a typical multi-decadal climate simulation at $1^{\circ}$ resolution, 25 nodes are on the low side.  However, this is the largest number of tasks supported by the $\tt f19$ grid, and we fixed the number of tasks across all grids for the purpose of comparing costs. We also show costs for 50 nodes, excluding $\tt f19$, to provide a more practical estimate for longer climate integrations. Other approaches for comparing costs across different grids and dycores, e.g., holding fixed the number of grid columns per task, are beyond the scope of this study.

>>>>

\begin{keypoints}
\item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two Arctic refined mesh configurations.
\item {\color{blue}{Quasi-uniform unstructured}} grids degrade the Greenland Ice Sheet mass balance compared to latitude-longitude grids, {\color{blue}{at the conventional $1^{\circ}$ resolution}}.
\item The refined Arctic meshes substantially improve the surface mass balance over {\color{blue}{conventional grid resolutions}}.
\end{keypoints}

@@

Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS). {\color{blue}{The finite-volume dynamical core uses structured, latitude-longitude grids, whereas the spectral-element dynamical core is built on unstructured meshes, permitting grid flexibility such as quasi-uniform grid spacing globally.}} The $1^{\circ}-2^{\circ}$ {\color{blue}{latitude-longitude and quasi-uniform unstructured}} grids systematically overestimate both accumulation and ablation over the GrIS. Of these $1^{\circ}-2^{\circ}$ grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids because they have more degrees of freedom to represent the GrIS. Two Arctic-refined meshes, with $1/4^{\circ}$ and $1/8^{\circ}$ refinement over Greenland, were developed for the spectral-element dynamical core and are documented here as newly supported configurations in CESM2.2. The Arctic meshes substantially improve the simulated clouds and precipitation rates in the Arctic. Over Greenland, these meshes skillfully represent accumulation and ablation processes, leading to a more realistic GrIS surface mass balance. As CESM is in the process of transitioning away from conventional latitude-longitude grids, these new Arctic-refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids, albeit at increased computational cost.

@@

The mass balance of Earth's big ice sheets is crucially important for understanding controls on global sea-level rise. However, the scale of the processes needed to represent ice sheet mass balance is challenging to resolve in conventional Earth System Models. This study evaluates the ability of different grids and atmospheric solvers (i.e, the dynamical core) in CESM2.2 to resolve the surface mass balance (SMB) of the Greenland Ice Sheet. We show that the ongoing transition away from latitude{\color{blue}{-}}longitude grids, towards quasi-uniform unstructured grids in CESM2.2 leads to a degradation of the simulated SMB. Two variable-resolution grids with enhanced resolution over Greenland are developed and incorporated into the release of CESM2.2, which substantially improves the SMB over the latitude longitude grids.
%[ enter your Plain Language Summary here or delete this section]

@@

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change. GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a, hereafter referred to as \textit{lat-lon} grids). Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain. With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic due to the polar singularity)}}. In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids {\color{blue}{(i.e., grids with enhanced resolution over a particular region)}}, to understand their impacts on the simulated Arctic climate. We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.

@@

Dycores built on lat-lon grids have some advantages over {\color{blue}{dycores built on}} unstructured grids. Lat-lon coordinate lines are orthogonal, and aligned with zonally symmetric circulations that characterize many large-scale features of Earth's atmosphere. \citeA{LJTN2010JAMES} has experimented with rotating lat-lon models such that their coordinate lines no longer align with an idealized, zonally balanced circulation. For the finite-volume lat-lon dycore considered in this paper (hereafter \textit{FV}), numerical errors were shown to be largest when the polar singularity is rotated into the baroclinic zone ($45^{\circ}$N latitude), generating spurious wave growth much earlier in the simulation than for other rotation angles. This illustrates the advantages of coordinate surfaces aligned with latitude bands, albeit an extreme example where the polar singularity and the polar filter are also contributing to the spurious wave growth. The unstructured grids all generate spurious baroclinic waves earlier in the simulations than the (unrotated) lat-lon models, although the unstructured model considered in this paper, the spectral-element dycore (hereafter \textit{SE}), holds a balanced zonal flow without spurious wave growth appreciably longer than the rotated FV experiments \cite{LJTN2010JAMES}. And unlike FV, the SE dycore has the same error characteristics regardless of how the grid is rotated.

The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which {\color{blue}{has a}} large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}. Unstructured grids support quasi-uniform grid spacing globally, and there is no pole problem (e.g., Figure~\ref{fig:uni-grids}c). {\color{blue}{This is in part why}} unstructured grids are becoming more common; their improved performance on massively parallel systems and lack of constraints on grid structure \cite{TTI1997JCP,PL2007JCP,WETAL2013GMD}. This increased grid flexibility allows for the adoption of variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement. In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).  However, local grid refinement comes at the cost of a smaller CFL-limited time step in the refined region; the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.

We emphasize that the pole problem is a distinctive feature of the dycore in atmospheric models. Polar filters do not directly interfere with the physical parameterizations, nor do they have any bearing on the surface models; e.g., the land model can take full advantage of the greater number of grid cells in polar regions on lat-lon grids. This is particularly relevant for the surface mass balance {\color{blue}{(\textit{SMB}; the integrated sum of precipitation and runoff)}} of the Greenland Ice Sheet, which relies on hydrological processes represented in the land model.

The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales {\color{blue}\cite{Fyke_EA_2018_Rev_Geophys}} that are difficult to represent in GCMs \cite{P2010CC}. GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation. The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias {\color{blue}{in the interior}} \cite{P2000GPC,VETAL2018TC}. GrIS ablation areas {\color{blue}{(marginal regions where the annual SMB is negative)}} are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation. These narrow ablation zones are not fully resolved in low-resolution GCMs, and may further degrade the simulated SMB. {\color{blue}{More recently, GCMs such as CESM and the NASA Goddard Institute for Space Studies GCM \cite{AETAL2019JGR}, have implemented an elevation class downscaling scheme for computing the SMB. The downscaling helps to resolve these narrow ablation zones in GCMs \cite{SETAL2019TC}, but large SMB biases remain.}} For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.

@@

In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} used the VR capabilities of the SE dycore in CESM, generating {\color{blue}{grids}} where Greenland is represented with {\color{blue}{up to}} $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution. The simulated SMB compared favorably to the SMB from RCMs and observations. The VR approach is therefore emerging as a powerful tool for simulating and understanding the GrIS and its response to different forcing scenarios.

@@

{\color{blue}{CESM2.2 is a CMIP6-class \cite<Coupled Model Intercomparison Project Phase 6;>{CMIP6} Earth System Model maintained by the National Center for Atmosphere Research. CESM2.2 contains sub-component models for the atmosphere, land, ocean, sea-ice, and land-ice, and can be configured to run with varying degrees of complexity. All simulations described in this study use an identical transient 1979-1998 Atmospheric Model Intercomparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}. In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2. The land-ice component is not dynamically active in $\tt FHIST$. However, the surface mass balance is computed by the land model before being passed to the land-ice component; $\tt FHIST$ includes the necessary functionality to simulate the SMB of the Greenland Ice Sheet.}}

@@

The FV dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}. The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}. Hyperviscous damping is applied to the divergent modes, and is increased in the top few layers (referred to as a \textit{sponge layer}) to prevent undesirable interactions with the model top, such as wave reflection \cite{L2011IJHPC}. A polar filter damps computational instability due to the convergence of meridians, permitting a longer time step. It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the meridional direction \cite{ST1995GEOS}. The form of the filter is designed to slow down the propagation of large zonal wave-numbers to satisfy the CFL condition of the shortest resolved wave at some reference latitude.

@@

The SE dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}. The computational domain is a cubed-sphere grid ({\color{blue}{obtained by projecting each face of a cube onto a sphere}}) tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}). Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre quadrature points. This results in 16 nodal points per element, with 12 of the points lying on the (shared) element boundary. Communication between elements uses the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries to reconcile overlapping nodal values and produce a continuous global basis set.

@@

{\color{blue}{SE is a next-generation dycore, and is less mature than the FV dycore due to its shorter history.}} In CESM2.2, the SE numerics have been enhanced {\color{blue}{relative to CESM2.0}} to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}. {\color{blue}{Future versions of CESM will likely include further optimizations and enhancements to the (more junior) SE dycore.}}

The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over {\color{blue}{quasi-uniform resolution grids}}. First, as the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. Second, the topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{LetAl2015GMD} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.

When using the SE dycore with quasi-uniform grid spacing, the SE tracer transport scheme is replaced with the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}. Atmospheric tracers have large, nearly discontinuous horizontal gradients that are difficult to represent with spectral methods, which are prone to oscillatory ``Gibbs-ringing" errors \cite{RW1990QJR}. CSLAM has improved tracer property preservation and accelerated multi-tracer transport. It uses a separate grid from the spectral-element dynamics, dividing each element into $3\times3$ control volumes with quasi-equal area. The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the original SE dycore in which the physics are evaluated at Gauss-Lobatto-Legendre points \cite{HL2018MWR}. {\color{blue}{CSLAM advection is not available in the VR configuration, which instead uses the standard SE tracer transport scheme with the physics evaluated at Gauss-Lobatto-Legendre points.}}

@@

Three VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic and the United States (Figure~\ref{fig:vr-grids}). This paper serves as the official documentation of these grids. The VR meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region ({\color{blue}{when using degrees to describe the resolution of unstructured grids, we refer to degrees at the equator}}). The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution. The $\tt CONUS$ grid contains $1/8^{\circ}$ refinement over the United States, and $1^{\circ}$ everywhere else. The $\tt CONUS$ grid is not discussed any further in this paper; see \citeA{PETAL2020BAMS} for simulations with the $\tt CONUS$ grid.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & {\color{blue}{cost(25 nodes)}} & {\color{blue}{cost(50 nodes)}} \\ 
   \hline
   $\tt f19$ & FV & 278 & - &1800 & {\color{blue}{436.66}} & {\color{blue}{-}} \\
   $\tt f09$ & FV & 139 & - &1800 & {\color{blue}{1534.57}} & {\color{blue}{2024.24}} \\
   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 & {\color{blue}{1497.26}} & {\color{blue}{1683.97}} \\
   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 & {\color{blue}{1890.48}} & {\color{blue}{2090.43}} \\
   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 & {\color{blue}{-}} & {\color{blue}{-}} \\
   $\tt Arctic$ & SE & 111 & 28 & 450 & {\color{blue}{15947.41}} & {\color{blue}{16675.45}} \\
   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 & {\color{blue}{40305.03}} & {\color{blue}{41036.67}} \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{fine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step. FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection. {\color{blue}{The FV dycore uses lat-lon grids, whereas the SE and SE-CSLAM dycores run on unstructured grids.}} We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$. {\color{blue}{The last columns provide the computational costs in core hours per simulated year (CHPSY). The costs are from single-month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}}}
 \label{tbl:table1}
 \end{table*}

The accuracy of the simulated surface mass balance is expected to be sensitive to grid resolution. Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids, {\color{blue}{as well as two grids pertaining to the Regional Atmospheric Climate Model \cite<RACMO;>{NETAL2018TC,NETAL2019SCIENCE}, which are used for validation (Table~\ref{tbl:table2})}}. The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160$ km, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6$ km, similar to the $11$ km grid spacing of the RACMO2.3 grid. The $\tt ne30pg3$ grid has an average $\Delta x=111.2$ km, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60$ km. Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs (Table~\ref{tbl:table1}), the convergence of meridians on the FV grid enhances the resolution over the GrIS. The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8$ km.

@@

\subsection{{\color{blue}{Computational costs}}}

{\color{blue}{The last columns of Table~\ref{tbl:table1} provide cost estimates for the different grids and dycores. The costs, expressed as core hours per simulated year (CHPSY), are taken from single-month runs of $\tt FHIST$, with no i/o, and using 25 nodes (900 tasks) and 50 nodes (1800 tasks) on the Cheyenne supercomputer \cite{cheyenne}. For a typical multi-decadal climate simulation at $1^{\circ}$ resolution, 25 nodes are on the low side.  However, this is the largest number of tasks supported by the $\tt f19$ grid, and we fixed the number of tasks across all grids for the purpose of comparing costs. We also show costs for 50 nodes, excluding $\tt f19$, to provide a more practical estimate for longer climate integrations. Other approaches for comparing costs across different grids and dycores, e.g., holding fixed the number of grid columns per task, are beyond the scope of this study.
